---
  title: "Machine Learning - Project to quantify how well personal activities are done"
  output: md_document
---
  
*Synopsis: This documet analyses the training dataset and predicts the category of the quality of exercise in the test dataset.
People commonly measure the amount of activiteies they perform, but seldom measure how well they do it. 
The aim of the current project is to analys the data from rom accelerometers on the belt, forearm, arm, and dumbell of 6 participants
and predict the category of the test data*

##Data Preprocessing 

```{r}
##The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har

#Create sub directory in current working directory if it doesn't exist
if(!file.exists("./data")){dir.create("./data")}
trainingFileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testFileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# Download File if not dowloaded already
if(!file.exists("./data/devise.csv")){ download.file(trainingFileUrl,destfile="./data/devise.csv") }
if(!file.exists("./data/test.csv")){ download.file(testFileUrl,destfile="./data/test.csv") }

# Unzip and store the data files
#unzip(zipfile="./data/storm.csv.bz2",exdir="./data")
path_rf <- file.path("./data")
```
Read the downloaded file
```{r, cache=TRUE}

# Read Files (into Data Frame)test and training
deviseData  <- read.csv(file.path(path_rf, "devise.csv" ))
dim(deviseData)


testData  <- read.csv(file.path(path_rf, "test.csv" ))
dim(testData)

```

Include libraries

```{r}
library(ggplot2); library(gdata) ; library(caret) ; library(randomForest)

set.seed(12345)
#From the training data create test and training data

inTrain <- createDataPartition(y=deviseData$classe,p=0.75, list=FALSE)

deviseTrain <- deviseData [inTrain,]
deviseTest <- deviseData [-inTrain,]

summary(deviseTrain)



# Plot graph to visually look at some key variables 
qplot(pitch_forearm , data=deviseTrain, geom="density", colour=classe, 
   main="Distribution Pitch Forearm", xlab="Pitch Forearm", 
   ylab="Density")

# Plot graph to visually look at some key variables 
qplot(pitch_arm , data=deviseTrain, geom="density", colour=classe, 
   main="Distribution Pitch Arm", xlab="Pitch Arm", 
   ylab="Density")

# Plot graph to visually look at some key variables 
qplot(pitch_belt , data=deviseTrain, geom="density", colour=classe, 
   main="Distribution Pitch Belt", xlab="Pitch Belt", 
   ylab="Density")

```

#Results:

Fit the model in test Data and calculate the training data

```{r}
#From the summary of the training data it is ovserved that first sevel columns are accessory data and not related to the activity output
#Also many of the data elements have a large number (14,408) of invalid data. all these are removed from the 
#Select the names of feature variables.

predCol <- matchcols(deviseTrain[,-(1:7)], without=c
                      ("skewness","kurtosis","max","min", "amplitude", "var_","stddev_","avg_", 
                      "classe"), method="or")

#Create the dataframe with 
predVars <- deviseTrain[,predCol]

#create the model with Random Forest 
modFit <- randomForest(x= predVars,y= deviseTrain$classe)

modFit

# Apply the model on the test data separated out of training data
test <- deviseTest[,predCol]
pred <- predict(modFit,test)

table(pred,deviseTest$classe)

```

**Conclusion: **
  
  *Random Forest has been used because of inherent accuracy of the model*
  
```{r}  
# Fitting the model on the supplied test Data  
testFinal <- testData[,predCol]
pred <- predict(modFit,testFinal)
pred
```
  
